---
title: "Week01 Practice"
output: html_notebook
---

Okay, so this will be a practice notebook for first week.

```{r}
library(tidyverse)
```


```{r}
data(mpg)
```

Dataset description:
- manufacturer: manufacturer name

- model: model name

- displ: engine displacement, in litres

- year:year of manufacture

- cyl: number of cylinders

- trans: type of transmission

- drv: the type of drive train, where f = front-wheel drive, r = rear wheel drive, 4 = 4wd

- cty: city miles per gallon

- hwy: highway miles per gallon

- fl: fuel type

- class: "type" of car


Dimensions:
```{r}
dim(mpg)
```

Head:
```{r}
head(mpg, 5)
```


Summary: 
```{r}
summary(mpg)
```
```{r}
glimpse(mpg)
```

```{r}
ggplot(data=mpg) +
  geom_point(mapping = aes(x = displ, y= cyl, color = class)) 
```

```{r}
ggplot(data=mpg) +
  geom_point(mapping = aes(x = displ, y= hwy, fill = class), shape=21, color='black') 
```
```{r}
ggplot(data=mpg) +
  geom_point(mapping = aes(x = displ, y= displ, color = displ < 5), shape=21) +
  facet_wrap(~ class, ncol = 3)
```
```{r}
ggplot(data=mpg) +
  geom_point(mapping = aes(x = displ, y= hwy), shape=21) +
  facet_grid(cty ~ .)
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color= class)) +
  facet_grid(drv ~ .)
```
```{r}
mpg$cylc = as.character(mpg$cyl)

ggplot(data=mpg) +
  geom_point(mapping = aes(x = displ, y= hwy, color = class, shape = cylc)) 
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class)) +
  facet_grid(. ~ cyl)
```
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

## Geometric objects

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth() + 
  geom_point(mapping = aes(color = class))
```

```{r}
ggplot(data = mpg) +
  geom_bar(mapping = aes(x=class))
```

```{r}
ggplot(data=mpg) +
  geom_histogram(mapping=aes(x=hwy))
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()
```


```{r}
ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth(se = FALSE, color = 'black')
```
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth(se = FALSE, color = 'blue', mapping = aes(group = drv))
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) +
  geom_point() +
  geom_smooth(se = FALSE)
```
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(color = drv)) +
  geom_smooth(se = FALSE)
```
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(color = drv)) +
  geom_smooth(mapping = aes(linetype = drv), se = FALSE)
```
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(fill = drv, stroke=2), size=3, shape=21, color='white')
```
## Statistical transformations

```{r}
data(diamonds)
str(diamonds)
```
```{r}
head(diamonds, 5)
```
```{r}
ggplot(data=diamonds) +
  geom_bar(mapping=aes(x=cut))
```
```{r}
ggplot(data = diamonds) + 
  stat_count(mapping = aes(x = cut))
```
```{r}
ggplot(data = diamonds) + 
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.min = min,
    fun.max = max,
    fun = median
  )
```
```{r}
ggplot(data = diamonds) +
  geom_pointrange(
    mapping = aes(x = cut, y = depth),
    stat = 'summary',
    fun.min = min,
    fun.max = max,
    fun = median
  )
```
```{r}
ggplot(data=diamonds) + 
  geom_col(mapping=aes(x=cut, y=depth))
```
## Position arguments 

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "fill")
```
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge")
```
### Excercises
 
```{r}
 ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point(position = 'jitter')
```
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_count(position ='jitter')
```
```{r}
ggplot(data=mpg) +
  geom_boxplot(mapping=aes(y=hwy))
```
## Coordinate systems

### Excercises


```{r}
bar <- ggplot(data = diamonds) + 
  geom_bar(
    mapping = aes(x = cut, fill = cut), 
    show.legend = FALSE,
    width = 1
  ) + 
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)

bar + coord_flip()
bar + coord_polar()
```

```{r}
bar <- ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "fill") +
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)

bar + coord_flip()
bar + coord_polar()

```
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline(slope = 1, intercept = 0) +
  coord_fixed()
```




