---
title: "Week01 Practice"
output: html_notebook
---

Okay, so this will be a practice notebook for first week.

```{r}
library(tidyverse)
```

# Data Visualisations

```{r}
data(mpg)
```

Dataset description:
- manufacturer: manufacturer name

- model: model name

- displ: engine displacement, in litres

- year:year of manufacture

- cyl: number of cylinders

- trans: type of transmission

- drv: the type of drive train, where f = front-wheel drive, r = rear wheel drive, 4 = 4wd

- cty: city miles per gallon

- hwy: highway miles per gallon

- fl: fuel type

- class: "type" of car


Dimensions:
```{r}
dim(mpg)
```

Head:
```{r}
head(mpg, 5)
```


Summary: 
```{r}
summary(mpg)
```
```{r}
glimpse(mpg)
```

```{r}
ggplot(data=mpg) +
  geom_point(mapping = aes(x = displ, y= cyl, color = class)) 
```

```{r}
ggplot(data=mpg) +
  geom_point(mapping = aes(x = displ, y= hwy, fill = class), shape=21, color='black') 
```
```{r}
ggplot(data=mpg) +
  geom_point(mapping = aes(x = displ, y= displ, color = displ < 5), shape=21) +
  facet_wrap(~ class, ncol = 3)
```
```{r}
ggplot(data=mpg) +
  geom_point(mapping = aes(x = displ, y= hwy), shape=21) +
  facet_grid(cty ~ .)
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color= class)) +
  facet_grid(drv ~ .)
```
```{r}
mpg$cylc = as.character(mpg$cyl)

ggplot(data=mpg) +
  geom_point(mapping = aes(x = displ, y= hwy, color = class, shape = cylc)) 
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class)) +
  facet_grid(. ~ cyl)
```
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

## Geometric objects

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth() + 
  geom_point(mapping = aes(color = class))
```

```{r}
ggplot(data = mpg) +
  geom_bar(mapping = aes(x=class))
```

```{r}
ggplot(data=mpg) +
  geom_histogram(mapping=aes(x=hwy))
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()
```


```{r}
ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth(se = FALSE, color = 'black')
```
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point() +
  geom_smooth(se = FALSE, color = 'blue', mapping = aes(group = drv))
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) +
  geom_point() +
  geom_smooth(se = FALSE)
```
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(color = drv)) +
  geom_smooth(se = FALSE)
```
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(color = drv)) +
  geom_smooth(mapping = aes(linetype = drv), se = FALSE)
```
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(fill = drv, stroke=2), size=3, shape=21, color='white')
```
## Statistical transformations

```{r}
data(diamonds)
str(diamonds)
```
```{r}
head(diamonds, 5)
```
```{r}
ggplot(data=diamonds) +
  geom_bar(mapping=aes(x=cut))
```
```{r}
ggplot(data = diamonds) + 
  stat_count(mapping = aes(x = cut))
```
```{r}
ggplot(data = diamonds) + 
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.min = min,
    fun.max = max,
    fun = median
  )
```
```{r}
ggplot(data = diamonds) +
  geom_pointrange(
    mapping = aes(x = cut, y = depth),
    stat = 'summary',
    fun.min = min,
    fun.max = max,
    fun = median
  )
```
```{r}
ggplot(data=diamonds) + 
  geom_col(mapping=aes(x=cut, y=depth))
```
## Position arguments 

```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "fill")
```
```{r}
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "dodge")
```
### Excercises
 
```{r}
 ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point(position = 'jitter')
```
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_count(position ='jitter')
```
```{r}
ggplot(data=mpg) +
  geom_boxplot(mapping=aes(y=hwy))
```
## Coordinate systems

### Excercises


```{r}
bar <- ggplot(data = diamonds) + 
  geom_bar(
    mapping = aes(x = cut, fill = cut), 
    show.legend = FALSE,
    width = 1
  ) + 
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)

bar + coord_flip()
bar + coord_polar()
```

```{r}
bar <- ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity), position = "fill") +
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)

bar + coord_flip()
bar + coord_polar()

```
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline(slope = 1, intercept = 0) +
  coord_fixed()
```
# Data Transformation

```{r}
library(nycflights13)
data(flights)
```

```{r}
str(flights)
```

## Filtering

```{r}
(flights.jfk <- dplyr::filter(flights, origin == "JFK"))
```

```{r}
# How many flights have a missing dep_time? What other variables are missing? What might these rows represent?
sum(is.na(flights$dep_time))
```

## Rearranging

```{r}
arrange(flights)
```
```{r}
rename(flights, tail_num = tailnum)
```

```{r}
select(flights, time_hour, air_time, everything())
```

## Mutating

```{r}
flights_sml <- select(flights, 
  year:day, 
  ends_with("delay"), 
  distance, 
  air_time
)
```

```{r}
mutate(flights_sml,
  gain = dep_delay - arr_delay,
  speed = distance / air_time * 60
)
```
```{r}
mutate(flights_sml,
  gain = dep_delay - arr_delay,
  hours = air_time / 60,
  gain_per_hour = gain / hours
)
```
```{r}
transmute(flights,
  gain = dep_delay - arr_delay,
  hours = air_time / 60,
  gain_per_hour = gain / hours
)
```
```{r}
transmute(flights,
  dep_time,
  hour = dep_time %/% 100,
  minute = dep_time %% 100
)
```
### Excercises

Currently `dep_time` and `sched_dep_time` are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.

```{r}
select(flights, dep_time, sched_dep_time) %>%
  mutate(dep_hour = dep_time %/% 100, 
         dep_minutes = dep_time %% 100,
         sched_dep_hour = sched_dep_time %/% 100, 
         sched_dep_minutes = sched_dep_time %% 100,
         dep_time = dep_hour * 60 + dep_minutes,
         sched_dep_time = sched_dep_hour * 60 + sched_dep_minutes)
```

Compare `air_time` with `arr_time` - `dep_time`. What do you expect to see? What do you see? What do you need to do to fix it?
```{r}
select(flights, arr_time, dep_time, air_time)
```

```{r}
select(flights, air_time, arr_time, dep_time) %>%
  mutate(c_air_time = arr_time - dep_time) %>% 
  select(c_air_time, everything())
```
```{r}
select(flights, air_time, arr_time, dep_time) %>%
  mutate(arr_time_1 = ((arr_time %/% 100) * 60 + (arr_time %% 100)) %% 1440,
         dep_time_1 = ((dep_time %/% 100) * 60 + (dep_time %% 100)) %% 1440,
         cc_air_time = (arr_time_1 - dep_time_1) - air_time)
```
## Summarise

```{r}
summarise(flights, delay = mean(dep_delay, na.rm = TRUE))
```
```{r}
dplyr::group_by(flights, carrier, origin) %>%
  summarize(delay = mean(dep_delay, na.rm = TRUE), c = n())
```
```{r}
dplyr::group_by(flights, carrier) %>%
  summarize(delay = mean(dep_delay, na.rm = TRUE), c = n())
```
```{r}
by_dest <- group_by(flights, dest)
delay <- summarise(by_dest,
  count = n(),
  dist = mean(distance, na.rm = TRUE),
  delay = mean(arr_delay, na.rm = TRUE)
)
#> `summarise()` ungrouping output (override with `.groups` argument)
delay <- filter(delay, count > 20, dest != "HNL")

# It looks like delays increase with distance up to ~750 miles 
# and then decrease. Maybe as flights get longer there's more 
# ability to make up delays in the air?
ggplot(data = delay, mapping = aes(x = dist, y = delay)) +
  geom_point(aes(size = count), alpha = 1/3) +
  geom_smooth(se = FALSE)
#> `geom_smooth()` using method = 'loess' and formula 'y ~ x'
```
```{r}
not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))

not_cancelled %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay))
```
```{r}
delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(
    delay = mean(arr_delay),
    n = n()
  )
#> `summarise()` ungrouping output (override with `.groups` argument)
ggplot(data = delays, mapping = aes(x = n, y=delay)) + 
  geom_point(alpha=.3)
```
```{r}
delays %>% 
  filter(n > 25) %>% 
  ggplot(mapping = aes(x = n, y = delay)) + 
    geom_point(alpha = 1/10)

```
```{r}
# Convert to a tibble so it prints nicely
batting <- as_tibble(Lahman::Batting)

batters <- batting %>% 
  group_by(playerID) %>% 
  summarise(
    ba = sum(H, na.rm = TRUE) / sum(AB, na.rm = TRUE),
    ab = sum(AB, na.rm = TRUE)
  )
#> `summarise()` ungrouping output (override with `.groups` argument)

batters %>% 
  filter(ab > 100) %>% 
  ggplot(mapping = aes(x = ab, y = ba)) +
    geom_point() + 
    geom_smooth(se = FALSE)
#> `geom_smooth()` using method = 'gam' and formula 'y ~ s(x, bs = "cs")'
```
```{r}
not_cancelled %>% 
  group_by(year, month, day) %>% 
  mutate(r = min_rank(desc(dep_time))) %>% 
  filter(r %in% range(r))
```

# EDA

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut))
```

```{r}
diamonds %>% 
  count(cut)
```
```{r}
ggplot(data=diamonds) +
  geom_histogram(mapping = aes(x = carat), binwidth = .1)
```





