---
title: "R Notebook"
output: html_notebook
---
```{r}
library(tidyverse)
```

# The data set
```{r}
file_url <- "https://covid.ourworldindata.org/data/owid-covid-data.csv"
data_set <- read.csv(file_url, 
                     header = TRUE,
                     check.names = FALSE,
                     stringsAsFactors = FALSE)
head(data_set)
```
```{r}
data_set <- data_set %>% 
  filter(!str_detect(iso_code, "^OWID_"))
```
# Task 0
```{r}
data_set$year <- substr(data_set$date, 1, 4)
plot_frame <- data_set %>% 
  select(new_cases, year) %>% 
  group_by(year) %>% 
  summarise(cases = sum(new_cases, na.rm = TRUE))
print(plot_frame)
```
Let’s define the visualization mapping in aes() and plot nothing, for the beginning:

```{r}
ggplot(data = plot_frame, 
       aes(x = year, y = cases))
```
Ok: ggplot2 recognizes my variables and their scales, but plots nothing. I need to add some geoms there for plot to materialize:
```{r}
ggplot(data = plot_frame, 
       aes(x = year, y = cases)) + 
  geom_point(size = 1.5, color = "darkred") + 
  geom_path(group = 1, size = .5, color = "darkred")
```
I could have added a different geom:
```{r}
ggplot(data = plot_frame, 
       aes(x = year, y = cases)) + 
  geom_bar(width = .5, stat = "identity", fill = "darkred", color = "black")
```
Now, what about total_cases per month and per year? I need a new plot_frame for this:
```{r}
data_set$year <- substr(data_set$date, 1, 4)
data_set$month <- as.character(substr(data_set$date, 6, 7))
plot_frame <- data_set %>% 
  select(total_cases, year, month) %>% 
  group_by(year, month) %>% 
  summarise(cases = sum(total_cases, na.rm = TRUE))
print(plot_frame)
```
```{r}
ggplot(data = plot_frame, 
       aes(x = month, 
           y = cases,
           fill = year, 
           color = year)) + 
  geom_bar(width = .5,
           stat = "identity", 
           position = position_dodge(width = 0.5))
```
Or I can do facet_wrap:
```{r}
ggplot(data = plot_frame, 
       aes(x = month, 
           y = cases,
           fill = year, 
           color = year)) + 
  geom_bar(width = .5,
           stat = "identity") + 
  facet_wrap(~year, nrow = 3)
```
Now I can begin using theme() in order to style my chart.
```{r}
ggplot(data = plot_frame, 
       aes(x = month, 
           y = cases,
           fill = year, 
           color = year)) + 
  geom_bar(width = .5,
           stat = "identity") + 
  facet_wrap(~year, nrow = 3) + 
  theme_light() + 
  theme(legend.position = "top")
```
Even better:
```{r}
ggplot(data = plot_frame, 
       aes(x = month, 
           y = cases,
           fill = year, 
           color = year)) + 
  geom_bar(width = .5,
           stat = "identity") + 
  facet_wrap(~year, nrow = 3) + 
  theme_bw() + 
  theme(legend.position = "top") + 
  theme(strip.background = element_rect(fill = "white", color = "white")) + 
  theme(panel.border = element_blank())
```
If you want to avoid using the scientific notation on a scale (look at our y-axis), you can use the scales library:
```{r}
# install.packages('scales')
library(scales)
ggplot(data = plot_frame, 
       aes(x = month, 
           y = cases,
           fill = year, 
           color = year)) + 
  geom_bar(width = .5,
           stat = "identity") + 
  facet_wrap(~year, nrow = 3) + 
  scale_y_continuous(labels = comma) + 
  theme_bw() + 
  theme(legend.position = "top") + 
  theme(strip.background = element_rect(fill = "white", color = "white")) + 
  theme(panel.border = element_blank())
```
Let’s now produce a scatter plot of the - mean number of new_cases per day, per country - and the country’s population (taking the latest available data).

Population first:
```{r}
population_set <- data_set %>% 
  select(location, date, population) %>% 
  arrange(location, date) %>% 
  group_by(location) %>% 
  slice_tail()
population_set
```
Now, the average number of total_cases per country:
```{r}
cases_set <- data_set %>% 
  select(location, new_cases) %>% 
  group_by(location) %>% 
  summarise(mean_cases_per_day = mean(new_cases, na.rm = TRUE))
cases_set
```
Now I need to join cases_set to population_set by locations; if you are interested to learn about relational structures and join operations in dplyr in R, go read section 13 Relational data in R for Data Science.
```{r}
plot_frame <- left_join(population_set, 
                        cases_set,
                        by = "location")
plot_frame
```
Here goes our scatter plot:
```{r}
ggplot(data = plot_frame, 
       aes(x = population, y = mean_cases_per_day)) + 
  geom_point(size = 1, color = "darkred") + 
  geom_point(size = .75, color = "white") + 
  scale_y_continuous(labels = comma) + 
  scale_x_continuous(labels = comma) + 
  theme_bw() + 
  theme(panel.border = element_blank())
```
Maybe using log() scales?
```{r}
ggplot(data = plot_frame, 
       aes(x = log(population), 
           y = log(mean_cases_per_day))) + 
  geom_point(size = 1, color = "darkred") + 
  geom_point(size = .75, color = "white") + 
  scale_y_continuous(labels = comma) + 
  scale_x_continuous(labels = comma) + 
  theme_bw() + 
  theme(panel.border = element_blank())
```
We can place the best fitting linear model on the chart by geom_smooth(method="lm):
```{r}
ggplot(data = plot_frame, 
       aes(x = log(population), 
           y = log(mean_cases_per_day))) + 
  geom_smooth(size = .25, method = "lm") + 
  geom_point(size = 1, color = "darkred") + 
  geom_point(size = .75, color = "white") + 
  scale_y_continuous(labels = comma) + 
  scale_x_continuous(labels = comma) + 
  theme_bw() + 
  theme(panel.border = element_blank())
```
## Task 1.
Reuse my code to produce a chart showing the total number of new_cases per continent and per year.
```{r}
task1_set <- data_set %>%
  select(continent, date, new_cases) %>%
  mutate(year = substr(date, 1, 4)) %>%
  group_by(continent, year) %>%
  summarise(new_cases = sum(new_cases, na.rm = TRUE), count = n())
```
```{r}
ggplot(data=task1_set, mapping=aes(x=continent, y=new_cases, color=year, fill=year)) +
  geom_bar(width=.4, stat='identity', position = position_dodge(width = 0.5)) + 
  scale_y_continuous(labels = comma) 
```

## Task 2.
Reuse my code to produce a scatter plot of:
* the mean gdp_per_capita per country, and
* the country population_density.
```{r}
task2_set <- data_set %>% 
  select(gdp_per_capita, location, continent, population_density) %>%
  group_by(location, continent) %>%
  summarise(gdp_per_capita_mean=mean(gdp_per_capita, na.rm = TRUE), 
            population_density=mean(population_density, na.rm = TRUE))
```
```{r}
ggplot(data=task2_set,
       mapping=aes(x=gdp_per_capita_mean, 
                   y=log(population_density), shape=continent)) +
  geom_point()
```



